---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r}
activity_data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
### Step 1: Splitting the steps data into smaller vectors

The number of steps is measured in fixed intervals. Therefore, we have 288 observations daily. I use this information to create 61 smaller lists.

```{r}
steps_by_groups <- split(activity_data$steps, ceiling(seq_along(activity_data$steps)/288))
```

### Step 2: Calculating the sum in each smaller vector

I'm looping through each column and calculate the sum.

```{r}
step_sums <- sapply(steps_by_groups,sum)
```

### Step 3: Creating the histogram

I draw the histogram with ggplot2.

```{r}
library(ggplot2)
qplot(step_sums, geom="histogram") 
```

### Step 4: Calculating the median and the mean of the number of steps
```{r}
steps_mean <- mean(step_sums, na.rm = TRUE)
steps_median <- median(step_sums, na.rm = TRUE)
```

The mean is `r steps_mean` and the median is `r steps_median`.

## What is the average daily activity pattern?

### Step 1 & 2: Time series plot and the maximum average measure.

```{r}
library(lattice)
library(data.table)
steps_sums_as_table <- data.frame(steps_by_groups)
average_steps_per_interval_across_days <- rowMeans(steps_sums_as_table, na.rm = TRUE)
ts_data <- ts(average_steps_per_interval_across_days, start = 1, end = 288)
xyplot(ts_data, type = "l")
max_average_measure <- max(ts_data)
```

The highest number of steps are made on average in the `r max_average_measure`th interval of the day. 

## Imputing missing values

### Step 1: Calculating the total number of missing values in the dataset.

```{r}
number_of_nas <- sum(is.na(activity_data))
```

There are `r number_of_nas` NA values in the dataset.

### Step 2: Devising a strategy for filling in all the missing values in the dataset. 

The strategy is to replace all NAs with the average number of steps taken, averages across all days.

